Synthesizing Symmetric Lenses

### 2018-07-12 papers/SymmetricLenses/main.tex
Checkpoint -- finished polishing intro

%\newif\ifdraft\drafttrue  % set true to show comments
\newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\iffull\fullfalse   % set true for long version
\newif\ifappendices\appendicesfalse

@@

{\em Lenses} are programs that can be run both ``front to back'' and ``back
to front,'' allowing data to be transformed in two directions.  Since their
introduction by~\citet{Focal2005-long}, lenses have been extensively
studied and applied.  Recent work~\cite{optician} has also demonstrated how
techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize a very simple class of lenses---so-called {\em
  bijective lenses} over string data---given a pair of types (regular
expressions) and a small number of examples.

We show how to extend this synthesis algorithm to a much richer class of
lenses, which we call {\em simple symmetric lenses}, including both
bijective lenses and the more widely used ``asymmetric''
lenses~\cite{Focal2005-long}, as well as a rich subset of the full-blown
``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric
lenses are of independent theoretical interest, being the largest class of
symmetric lenses that do not rely on persistent internal state.

@@

\noindent
In the big-data world, similar information is often stored or transmitted in
different formats. For instance, electronic calendars come in iCalendar,
vCalendar, and h-event formats~\cite{calendar-formats}; US taxation data can be
transmitted via Tax XML (used by the US government for e-filing) or TXF (used by
TurboTax)~\cite{tax-formats}; and sewing machines can use embroidery files in
formats with names like PES, DST, and several others~\cite{embroidery-formats}.
\bcp{It would be nice to be able to give a quick example of a truly symmetric
  case---do any of these examples work for that?}\afm{h-event and vCalendar are
  truly symmetric. Even the fields related to location alone are truly
  symmetric (though it is subtle to see this).}

One convenient way to interconvert between these varied data formats is to
use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that can
transform data represented in a format $S$ to a format $T$, and vice versa,
providing ``round tripping'' guarantees that help ensure data is not lost or
corrupted as it is transformed back and forth between different
representations.

While domain-specific languages for writing lenses can facilitate building
principled data transformations, \textit{synthesizing} lenses 
can make the task even easier.  Indeed, lens languages offer excellent
targets for synthesis, since their specialized, sub-Turing-complete syntax
and expressive type systems drastically
reduce the space of possible programs, making search-based synthesis easier. 
%
Recent work~\cite{optician} has exploited this observation, demonstrating that
it is feasible to synthesize quite complex examples, albeit only for a simple
form of lenses---so-called bijective lenses. Specifically, given a pair of types
describing the intended source and target formats plus a small number of
examples of corresponding data, the bijective synthesis algorithm will find a
bijective lens that converts between data stored the provided formats. Often,
these algorithms can find an appropriate lens in less than a second without
requiring any examples.
%, with just types describing the source and target formats.
% \dpw{I would cite the ICFP paper (cite string:
%   maina+:quotient-synthesis) as to appear but I'm not sure how to do that
%   without possibly violating the spirit of double-blind reviewing. ie, if it
%   hasn't appeared yet, how could we know the content if we aren't the authors?
%   Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
%   2018 website lists the paper; I think it is okay to cite it. The reviewer can
%   assume we contacted the authors for a copy. We could ask Stephanie for a
%   ruling if we want to be sure. One of the principles of double blind as adopted
%   by SIGPLAN is not to reduce the quality of the paper; omitting this citation
%   would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
%   third person, even if that's a bit awkward, but otherwise citing it is OK.}
One reason that bijective lens synthesis is so effective, even relative to successful
synthesis projects in other domains, is that there are not very
many bijections between related data formats (unless the
same kind of data appears in the two formats many times). 
Hence, if the synthesis algorithm can
find any bijection at all, then it is very likely to find the one the
user wants.

However, there is a price for narrowing the synthesis domain so dramatically:
many synchronization tasks cannot by synthesized! Many real data formats are
almost, but not quite, bijections. More often two related data formats have some
data in common, but also contain some differences. For instance, two related
tables might share four of five columns, but each has a different fifth column.
Alternately, one ad hoc configuration file might include some file-specific meta
data, such as a date, a time or a reference number, whereas the other does not.
Indeed of the 48 benchmark problems described in Section \ref{sec:evaluation},
all of the benchmarks taken from Flash Fill~\cite{flashfill} and many of the
benchmarks that synchronize between two ad hoc file formats have this
characteristic, 13 benchmarks in total. Consequently, searching only for strict
bijections is often not applicable in practice.

A natural way to expand the scope of such tools is to consider a more
expressive language of lenses.  For example, \emph{symmetric
lenses}~\cite{symmetric-lenses} encompass a broad collection of bidirectional
transformations that include bijections as well as more flexibile transformations that
allow each side to retain information not present
in the other.  If $\ell$ is a symmetric lens between formats $S$ and
$T$, then $\ell$  can transform data from $S$ to $T$ or vice versa
equally well.  A user would typically 
run $\ell$ in the forwards direction (from $S$ to $T$) when a user updates
a string $s \in S$ to propagate those updates to $T$.
However, because data in format $S$ may contain some information not present in format $T$,
the $S \rightarrow T$ transformation takes
\emph{both} the edited string $s \in S$ and a string $t \in T$ as arguments.
Inuitively, the lens extracts the data ``shared'' between the two
formats from $s$
and uses $t$ to supply any missing information.
% required by $T$ that is missing
%as it generates an output $t'$.  If $s$ is missing information, the transformation should
%extract the missing components from $t$.

Initially, we hypothesized that extending the synthesis algorithms
of \citet{optician} to operate over symmetric rather than
bijective lenses
would be relatively straightforward:  Simply replace the bijective
combinators with symmetric ones 
and search using similar heuristics.  However, 
this na\"ive approach runs into two difficulties.

The first difficulty is pragmatic. Symmetric lenses usually operate
over three structures: the source $S$, the target $T$ and a complement
$C$ that contains the information not
present in either $S$ or $T$~\cite{symmetric-lenses}. Logistically, the complement must be stored
somewhere and managed. More significantly, complements complicate synthesis
specifications.
Users can no longer just give examples of source and target
pairings; instead they must also indicate how to manage the 
``hidden state.'' To avoid such complexities, we define a new variant
of symmetric lenses: \emph{simple symmetric lenses}. At the price
of some expressiveness, these new lenses do not require complements.
%only the usual source and target.
%
To characterize their expressive power,
we prove that simple symmetric lenses include all (and only)
those symmetric lenses that satisfy an intuitive property called
\emph{forgetfulness}.
%, which we define in Section~\ref{sec:relationship}.
Intuitively, simple symmetric lenses are the complete set of symmetric
lenses that do not require external ``memory'' to recover data from past
instances of $S$ or $T$ when making a round trip. They only need the most
recent instance. 
We justify this loss of expressiveness pragmatically. 
We extend Boomerang~\cite{Boomerang} with simple symmetric lenses and
successfully use them on a range of real-world applications, showing
in the process that simple symmetric lenses integrate smoothly with 
other advanced lenses in Boomerang, including quotient
lenses~\cite{quotientlenses} and matching lenses~\cite{matchinglenses}.

The second difficulty with synthesizing symmetric lenses is more
fundamental: While the number of bijective lenses between two
related formats is often tiny, the number of simple symmetric lenses is much larger. In
practice, the difference in cardinality means that when a search algorithm
returns the first bijective lens it finds, that lens has a good chance of being
the desired lens. In contrast, and to our initial surprise, experiments showed
that if a na\"ive search algorithm selects the first simple symmetric lens it
finds, it is typically \emph{not} the desired lens.  The flexibility
that such lenses provide in eliding information 
generates an explosion of possible transformations. To disambiguate
between the many possible lenses, we need new 
principles for identifying the ``likely'' ones, and we need new
algorithms guided by those principles to
search the space of program transformations.

To overcome this second difficulty, we turn to information theory 
to characterize the likely and unlikely lenses. 
We consider a ``likely'' lens to be one that propagates a lot of
information from source to 
target and vice versa. Conversely, an ``unlikely'' lens requires a
large amount of additional information to recover the target given a
source and vice versa.  
This categorization follows from the intuition
that users will typically want to synchronise formats with lots
of shared information, and so lenses that preserve 
more information are preferable.
To implement this
intuition, we turn to \emph{stochastic regular expressions}
(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of strings \`a la
regular expressions and a probability distribution over those
strings.  Using this information, we can calculate the shared
information content of two formats and our measure of
likeliness.

\begin{figure}
  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
  \vspace{-2ex}
  \caption{Schematic diagram for the simple symmetric lens synthesis algorithm.
    The user provides regular expressions \BRegex and \BRegexAlt and a set of
    examples \Examples{} as input. \Expand first converts \BRegex and \BRegexAlt
    to stochastic regular expressions \Regex and \RegexAlt with default
    probabilities. It then finds pairs of stochastic regular expressions
    equivalent to \Regex and \RegexAlt and iteratively proposes them to
    \GreedySynth. \GreedySynth finds a lens typed between the supplied SREs.
    When the algorithm determines it has found a likely lens, it returns it.}
  \label{fig:high-level-algorithm}
\end{figure}

With simple symmetric lenses and our SRE-based likeliness measure in
hand, we develop a new algorithm for synthesizing likely lenses.
At its core, the algorithm performs a type-directed search between the 
descriptions of the source and target formats, measuring success using
the likeliness measure.
However, regular expressions enjoy a rich algebraic structure, meaning
there are infinitely many regular expressions equivalent to the
original pair, and the lens returned by a type-directed search depends
upon the chosen representation. 
To tame this complexity, we divide the synthesis algorithm into two communicating
search procedures (Figure~\ref{fig:high-level-algorithm}), following
earlier work~\cite{optician}.
The first algorithm, \Expand, uses rewriting rules to iteratively
propose new pairs of stochastic regular expressions equivalent to the
original pair.
The second algorithm, \GreedySynth, takes a pair of such SREs as input
and uses a greedy, type-directed algoithm to find a simple symmetric
lens between them, returning the lens and its likelihood score
to \Expand.  The algorithm terminates when the likilihood score of the
best lens found so far compared with 
the complexity of newly generated SREs 
suggests that further searching is unlikely to be fruitful.

@@

We implemented the \GreedySynth and \Expand algorithms and evaluated their
effectiveness at synthesizing lenses on a benchmark suite of 48 problems,
including data cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014},
view maintenance tasks taken from Augeas~\cite{Augeas}, and ad hoc format
synchronization tasks taken from a variety of other locations\footnote{The
  majority of these format synchronization tasks were taken from the
  bidirectional programming literature.}.  Our tool is intended to be responsive
enough for interactive use by the programmer, who can control the search process
via examples and a ``tuning'' parameter that governs how much effort is spent
trying to improve on existing solutions.

In summary, our contributions are:\bcp{The section references do not proceed
  in order, so this doesn't work very well as an outline of the paper.
  Should we try to reorder the contributions?  Write a separate outline?
  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's nice
  when the contributions mirror the paper, I am not jarred when they don't.}
\begin{itemize}
\item We define {\em simple symmetric lenses}, with the
  the useful property that they do not maintain any hidden state
  (\S\ref{sec:overview}).  
  We present a collection of combinators for defining these lenses and
  describe how to 
  calculate their likelihoods (\S\ref{sec:ssl}). 
\item  We show that simple symmetric lenses
  fall between asymmetric lenses~\cite{Focal2005-long} and full
  symmetric lenses~\cite{symmetric-lenses} in expressiveness
  (\S\ref{sec:related}).  We further show that the class of simple
  lenses can be characterized semantically as the subset of full
  symmetric lenses for which a simple ``forgetfulness'' property holds.
\item We show how to synthesize simple symmetric lenses (including all
  asymmetric lenses as a special case), using a novel application of stochastic
  regular expressions to guide the search process (\S\ref{sec:synthesis}). 
  To support that effort, we show that the star semi-ring equivalences on
  regular 
  expressions remain valid when applied to stochastic regular
  expressions, preserving the probability distributions as well as the languages
  (\S\ref{sec:sre}).
\item We extend the Boomerang implementation with simple symmetric lenses,
  demonstrating that they integrate smoothly with other lens extensions such as
  quotient lenses~\cite{quotientlenses} and matching
  lenses~\cite{matchinglenses}.  We evaluate our implementation and the effects
  of various optimizations and the tuning parameter on a set of 48 lens
  synthesis benchmarks drawn from data cleaning, view maintenance, and file
  synchronization tasks.  We demonstrate that, with only modest interactive
  tuning, the system can synthesize simple symmetric lenses for all of the benchmarks in
  under 30 seconds (\S\ref{sec:evaluation}).
\end{itemize}
We close with a discussion of related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).

@@

While classical symmetric lenses
are technically more expressive than simple symmetric lenses, it is unclear
whether this additional expressiveness is useful in practice. In our judgement,
the complications they introduce into synthesis did not justify the potential
additional expressiveness: because each lens has a custom
complement, one can no longer specify the put functions through input/output
examples alone. One alternative would be to enrich specifications with edit
sequentials; another would be to specify the structure of complements explicitly
(though the latter would be somewhat akin to specifying the internal state of a
program). In either case, the complexity of the specifications increases.
%% addition, users would have to specify some auxilary 
%% to specify the behavior through the observable behavior of edit sequences.
%% Second, there are lenses that require arbitrarily long edit sequences to
%% differentiate between: consider the lens that keeps two formats equal for the
%% first $k$ non-identity edits, then keeps them completely unsynchronized for all
%% future edits.  This requires an edit sequence of $k$ edits to differentiate
%% between this lens and the identity lens.  Simple symmetric lenses avoid this
%% problem by disallowing such lenses: we can't keep track of the number of
%% non-identity edits performed unless there are explicit fields in the data
%% formats keeping track of them.  While we are reducing the expressivity of our
%% lenses with this restriction, we feel the benefits outweigh the costs.

>>>>

% \newif\ifdraft\drafttrue  % set true to show comments
\newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\iffull\fullfalse   % set true for long version
\newif\ifappendices\appendicesfalse

@@

% BCP: To make sure we don't leave comments on!
\bcp{XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX }

{\em Lenses} are programs that can be run both ``front to back'' and ``back
to front,'' allowing data and updates to be transformed in two directions.
Since their 
introduction by~\citet{Focal2005-long}, lenses have been extensively
studied and applied.  Recent work~\cite{optician} has also demonstrated how
techniques from {\em type-directed program synthesis} can be used to
efficiently synthesize a very simple class of lenses---so-called {\em
  bijective lenses} over string data---given a pair of types (regular
expressions) and a small number of examples.

We show how to extend this synthesis algorithm to a much wider class of
lenses, which we call {\em simple symmetric lenses}, including both
bijective lenses and the more widely used ``asymmetric''
lenses~\cite{Focal2005-long}, as well as a rich subset of the full-blown
``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric
lenses are of independent theoretical interest, being the largest class of
symmetric lenses that do not rely on persistent internal state.

@@

\noindent
In today's big-data world, similar information is sometimes
stored in different places and in different formats. For instance, electronic
calendars come in iCalendar, vCalendar, and h-event
formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax
XML (used by the US government for e-filing) or TXF (used by
TurboTax)~\cite{tax-formats}; and sewing machines can use embroidery files
in formats with names like PES, DST, and several
others~\cite{embroidery-formats}.
%
\bcp{I'm not satisfied with the embroidery example: It doesn't obviously
  correspond to a bidirectional situation (the sewing machine does not need
  to send back the embroidery file to anybody).  The tax example could be
  criticized on the same grounds.}
%
\bcp{Also, it would be nice to be able to give a quick example of a truly
  symmetric case---do any of these examples work for that?}\afm{h-event and
  vCalendar are truly symmetric. Even the fields related to location alone
  are truly symmetric (though it is subtle to see this).}

One convenient way to maintain consistency between these varied data formats
is to use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that
can transform updates to data represented in a format $S$ to another format
$T$ and vice versa, providing ``round tripping'' guarantees that help ensure
information is not lost or corrupted as it is transformed back and forth
between different representations.

While domain-specific languages for writing lenses can facilitate building
principled data transformations, \textit{synthesizing} lenses 
can make the task even easier.  Indeed, lens languages offer excellent
targets for synthesis: their specialized, sub-Turing-complete syntax
and expressive type systems drastically
reduce the space of possible programs, making search-based synthesis easier. 
%
Recent work~\cite{optician} has exploited this observation, demonstrating that
it is feasible to synthesize quite complex examples, albeit only for a
restricted class of lenses---so-called bijective lenses. Specifically, given 
source and target formats plus a small number of
examples of corresponding data, the bijective synthesis algorithm will find a
bijective lens matching the examples, if one exists. Often,
it can find an appropriate lens in less than a second with
no examples at all.
%, with just types describing the source and target formats.
% \dpw{I would cite the ICFP paper (cite string:
%   maina+:quotient-synthesis) as to appear but I'm not sure how to do that
%   without possibly violating the spirit of double-blind reviewing. ie, if it
%   hasn't appeared yet, how could we know the content if we aren't the authors?
%   Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
%   2018 website lists the paper; I think it is okay to cite it. The reviewer can
%   assume we contacted the authors for a copy. We could ask Stephanie for a
%   ruling if we want to be sure. One of the principles of double blind as adopted
%   by SIGPLAN is not to reduce the quality of the paper; omitting this citation
%   would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
%   third person, even if that's a bit awkward, but otherwise citing it is OK.}
One reason that bijective lens synthesis is so effective, even relative to
successful synthesis tools in other domains, is that there are typically not
very many bijections between related data formats. 
Hence, if the synthesis algorithm can
find any bijection at all, it is likely to be the right one.

However, many real-world data formats share just part of their information
content with other representations.  For instance, two related
database tables might share four out of five columns, but each might have a
different fifth column. 
Or, one ad-hoc system configuration file might include some format-specific meta
data, such as a date or a reference number, whereas the same configuration
file on another operating system does not. 
Indeed of the 48 benchmark problems described in Section \ref{sec:evaluation},
all of the benchmarks taken from Flash Fill~\cite{flashfill} and many of the
benchmarks that synchronize between two ad hoc file formats have this
characteristic, 13 benchmarks in total. Consequently, synthesizing only strict
bijections is often not useful in practice.

Could we extend synthesis tools to more kinds of lenses?  In particular,
what about \emph{symmetric lenses}~\cite{symmetric-lenses}, a class that
includes bijective lenses, ``asymmetric'' lenses where the transformation
from $S$ to $T$ can throw away information that must then be restored when
returning from $T$ to $S$, and even more flexibile transformations that
allow each side to retain information not present in the other?

\bcp{We use $S$ and $T$ here but $X$ and $Y$ below.}

% A symmetric lens between $S$ and $T$ represents a pair of functions, $\PutR$
% and $\PutL$.  The $\PutR$ component has type
% $S \times T \rightarrow T$---that is, it takes a ``new'' $S$ and an ``old''
% $T$ and builds a ``new'' $T$ using its first argument for the shared
% information and its second argument for information that is only available
% on the $T$ side.  Conversely, the $\PutL$ component has type
% $T \times S \rightarrow S$.

% If $\ell$ is a symmetric lens between formats $S$ and
% $T$, then $\ell$  can transform data from $S$ to $T$ or vice versa
% equally well.  A user would typically 
% run $\ell$ in the forwards direction (from $S$ to $T$) when a user updates
% a string $s \in S$ to propagate those updates to $T$.
% However, because data in format $S$ may contain some information not present in format $T$,
% the $S \rightarrow T$ transformation takes
% \emph{both} the edited string $s \in S$ and a string $t \in T$ as arguments.
% Inuitively, the lens extracts the data ``shared'' between the two
% formats from $s$
% and uses $t$ to supply any missing information.
%
% required by $T$ that is missing
%as it generates an output $t'$.  If $s$ is missing information, the
%transformation should 
%extract the missing components from $t$.

One might hope that extending the algorithms from \citet{optician} to
synthesize symmetric rather than bijective lenses would be relatively
straightforward: Simply replace the bijective combinators with symmetric
ones and search using similar heuristics.  However, this na\"ive approach
encounters two difficulties.

The first is pragmatic. ``Classical'' symmetric lenses as presented
by ~\citet{symmetric-lenses} operate
over three structures: the source $S$, the target $T$ and a complement
$C$ that contains the information not
present in either $S$ or $T$. Logistically, these complements must be stored
and managed somehow.  More fundamentially, complements complicate
synthesis {\em specifications}---instead of just giving single examples of
source and target 
pairings, users would have to give longer ``interaction sequences'' to 
show how a lens should behave. To avoid these complexities, we define a
restricted variant 
of symmetric lenses, which we call \emph{simple symmetric lenses}. 
%only the usual source and target.
%
%, which we define in Section~\ref{sec:relationship}.
Intuitively, simple symmetric lenses are the complete set of symmetric
lenses that do not require external ``memory'' to recover data from past
instances of $S$ or $T$ when making a round trip. They only need the most
recent instance.

To characterize their expressive power formally, we show that simple
symmetric lenses are exactly the symmetric lenses that satisfy an intuitive
property called \emph{forgetfulness}.  To check that they are expressive
enough in practice, we add simple symmetric lenses to the Boomerang
language~\cite{Boomerang} and apply them to a range of real-world
applications, showing in the process that simple symmetric lenses integrate
smoothly with other advanced lens features provided by Boomerang, including
quotient lenses~\cite{quotientlenses,maina+:quotient-synthesis} and matching
lenses~\cite{matchinglenses}.

However, we are still left with a second difficulty with synthesizing
symmetric lenses: While the number of bijective lenses between two related
formats is typically tiny, the number of simple symmetric lenses is
typically huge.  Thus, if a na\"ive search algorithm selects the first
simple symmetric lens it finds, this will generally \emph{not} be the one we
want.  We need a new principle for identifying ``more likely'' lenses and a
more sophisticated synthesis algorithm that uses this principles to search
the space more intelligently.

For these, we turn to information theory. 
We consider a ``likely'' lens to be one that propagates ``a lot'' of
information from source to 
target and vice versa. Conversely, an unlikely lens requires a
large amount of additional information to recover the target given a
source.  
This categorization follows from the intuition
that users will typically want to synchronise formats with lots
of shared information, and so lenses that propagate
more information are preferable to ones that propagate less.
%
This intuition is formalized using \emph{stochastic regular expressions}
(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of
strings and a probability distribution over those strings.  Using this
probability distribution, we can calculate the shared information content of
two formats\bcp{It's not obvious at this point what the {\em shared}
  information content has to do with anything.} and our measure of likelihood.

\begin{figure}
  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
  \vspace{-2ex}
  \caption{Schematic diagram for the simple symmetric lens synthesis algorithm.
    The user provides regular expressions \BRegex and \BRegexAlt and a set of
    examples \Examples{} as input. \Expand first converts \BRegex and \BRegexAlt
    to stochastic regular expressions \Regex and \RegexAlt with default
    probabilities. It then finds pairs of stochastic regular expressions
    equivalent to \Regex and \RegexAlt and iteratively proposes them to
    \GreedySynth. \GreedySynth finds a lens typed between the supplied SREs.
    When the algorithm finds a likely lens, it returns it.\bcp{Later: We
      could save significant space if we put the inputs and outputs on the
      sides instead of top and bottom}}
  \label{fig:high-level-algorithm}
\end{figure}

With simple symmetric lenses and our SRE-based likelihood measure in
hand, we develop a new algorithm for synthesizing likely lenses.
At its core, the algorithm performs a type-directed search between the 
descriptions of the source and target formats, measuring success using
the likelihood measure.
The interesting complications arise from the need to deal with
regular-expression equivalences.  There are infinitely many regular
expressions equivalent to a given one, and the lens returned by a
type-directed search will depend which of the infinitely many possible
representations are chosen for its source and target formats.  Moreover,
there may be {\em no} lens connecting the 
source and target formats selected by the user: in general, the synthesis
algorithm has to search to find equivalent ones for which a suitable lens
exists.  
To tame this complexity, we divide the synthesis algorithm into two
communicating search procedures (Figure~\ref{fig:high-level-algorithm}),
following~\citet{optician}. 
The first, \Expand, uses rewriting rules to iteratively
propose new pairs of stochastic regular expressions equivalent to the
original pair.
The second, \GreedySynth, takes a pair of SREs as input
and uses a greedy, type-directed algorithm to find a simple symmetric
lens between them, returning the lens and its likelihood score
to \Expand.  The whole synthesis algorithm terminates when comparing the
likilihood score of the best lens found so far with 
the complexity of newly generated SREs 
suggests that further searching is unlikely to be fruitful.

@@

We added this synthesis procedure to the Boomerang implementation and
evaluated its effectiveness on a benchmark suite of 48 problems, including
data-cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014},
view-maintenance tasks taken from Augeas~\cite{augeas2}, and ad-hoc
synchronization tasks from a variety of other sources (mostly from the
bidirectional programming literature).  The ``programmer'' can control the
search process via examples and a {\em tuning parameter} that governs how
much effort is spent trying to improve on existing solutions.

In summary, our contributions are:\bcp{The section references do not proceed
  in order, so this doesn't work very well as an outline of the paper.
  Should we try to reorder the contributions?  Write a separate outline?
  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's
  nice when the contributions mirror the paper, I am not jarred when they
  don't.}\bcp{But this way, the reader has no birds-eye view of the paper at
  this point...}
\begin{itemize}
%
\item We define {\em simple symmetric lenses}, a subclass of symmetric
lenses that do not maintain any hidden state (\S\ref{sec:overview}).  We
present a collection of combinators for building simple symmetric lenses and
describe how to calculate their likelihoods (\S\ref{sec:ssl}).
%
\item We show that simple symmetric lenses fall between asymmetric
lenses~\cite{Focal2005-long} and full symmetric
lenses~\cite{symmetric-lenses} in expressiveness.  We further show that the
class of simple lenses can be characterized semantically as the subset of
full symmetric lenses for which a simple ``forgetfulness'' property holds
(\S\ref{sec:related}).
%
\item We give an algorithm for synthesizing simple symmetric lenses, using a
novel application of stochastic regular expressions to guide the search
process (\S\ref{sec:synthesis})\bcp{``and prove it correct''???}.  Along the
way, we show that {\em star semi-ring equivalences} on regular expressions
remain valid when applied to stochastic regular expressions, preserving
probability distributions as well as languages (\S\ref{sec:sre}).
%
\item We extend the Boomerang implementation with simple symmetric lenses,
demonstrating that they integrate smoothly with other lens extensions such
as quotient lenses~\cite{quotientlenses} and matching
lenses~\cite{matchinglenses}.  We evaluate our implementation and the
effects of various optimizations and the tuning parameter on a set of 48
lens synthesis benchmarks drawn from data cleaning, view maintenance, and
file synchronization tasks.  With only modest interactive tuning, the system
can synthesize simple symmetric lenses for all of the benchmarks in under 30
seconds (\S\ref{sec:evaluation}).
\end{itemize}
We close with related work (\S\ref{sec:related}) and
concluding thoughts (\S\ref{sec:conc}).

@@

While classical symmetric lenses
are technically more expressive than simple symmetric lenses, it is unclear
whether this additional expressiveness is useful in practice.\bcp{That's a
  total cop-out! We have much more to say about this point, and we {\em
    should} have even more to say than we do.} In our judgement,
the complications they introduce into synthesis did not justify the potential
additional expressiveness: because each lens has a custom
complement, one can no longer specify the put functions through input/output
examples alone. One alternative would be to enrich specifications with edit
sequentials; another would be to specify the structure of complements explicitly
(though the latter would be somewhat akin to specifying the internal state of a
program). In either case, the complexity of the specifications increases.
%% addition, users would have to specify some auxilary 
%% to specify the behavior through the observable behavior of edit sequences.
%% Second, there are lenses that require arbitrarily long edit sequences to
%% differentiate between: consider the lens that keeps two formats equal for the
%% first $k$ non-identity edits, then keeps them completely unsynchronized for all
%% future edits.  This requires an edit sequence of $k$ edits to differentiate
%% between this lens and the identity lens.  Simple symmetric lenses avoid this
%% problem by disallowing such lenses: we can't keep track of the number of
%% non-identity edits performed unless there are explicit fields in the data
%% formats keeping track of them.  While we are reducing the expressivity of our
%% lenses with this restriction, we feel the benefits outweigh the costs.
